\chapter{Command Summary} \label{appx:command}

\bigskip \begin{center}
Database Editing Commands (page~\pageref{cmd:dbedit})
\end{center}

\cmdsum{\cmd{TITLE}
} {
sets the title to be written to the output database.
}

\bigskip \begin{center}
Variable Selection Commands (page~\pageref{cmd:varsel})
\end{center}

\cmdsum{\cmd{SAVE}
   \param{variable$_{1}$}, \param{variable$_{2}$}, \ldots\
   or \param{option$_{1}$}, \param{option$_{2}$}, \ldots\
} {
transfers variables from the input database to the output database.
}

\cmdsum{\cmd{DELETE}
   \param{variable$_{1}$}, \param{variable$_{2}$}, \ldots\
} {
marks an assigned variable as a temporary variable that will not be
written to the output database.
}

\bigskip \begin{center}
Time Step Selection Commands (page~\pageref{cmd:timesel})
\end{center}

\input{../include/timesum}

\bigskip \begin{center}
Mesh Limiting Commands (page~\pageref{cmd:meshlimit})
\end{center}

\cmdsum{\cmd{ZOOM}
   \param{xmin}, \param{xmax}, \param{ymin}, \param{ymax},
      \param{zmin}, \param{zmax} [\cmd{OUTSIDE}]
} {
sets the limits of the mesh to be written to the output database.
}

\cmdsum{\cmd{VISIBLE}
   [\cmd{ADD} or \cmd{DELETE},]
   \param{block\_id$_{1}$}, \param{block\_id$_{2}$}, \ldots\
} {
limits the element blocks to be written to the output database.
}

\cmdsum{\cmd{FILTER ELEMENT}
  \param{variable} \param{lt$|$le$|$ge$|$ne$|$gt$|$ge} \param{value} \cmd{TIME} \param{db\_time}
} {
will delete all elements that satisfy the specified condition. 
}

\cmdsum{\cmd{REMOVE ELEMENT} 
  [\cmd{GLOBAL} or \cmd{LOCAL}] \param{id$_1$} \param{id$_2$} \ldots\ \param{id$_n$}
} {
  will remove the elements with the specified global or local id(s).
}

\bigskip \begin{center}
Element Block Selection Commands (page~\pageref{cmd:blocksel})
\end{center}

\cmdsum{\cmd{BLOCKS}
   [\cmd{ADD} or \cmd{DELETE},]
   \param{block\_id$_{1}$}, \param{block\_id$_{2}$}, \ldots\
} {
selects the element blocks which have defined values for all following
equations.
}

\cmdsum{\cmd{MATERIAL}
   [\cmd{ADD} or \cmd{DELETE},]
   \param{block\_id$_{1}$}, \param{block\_id$_{2}$}, \ldots\
} {
is exactly equivalent to a \cmd{BLOCKS} command.
}

\newpage %%%
\bigskip \begin{center}
Information and Termination Commands (page~\pageref{cmd:infoterm})
\end{center}

\cmdsum{\cmd{SHOW}
   \param{command}
} {
displays the settings of parameters relevant to the \param{command}.
}

\cmdsum{\cmd{LIST}
   \param{option}
} {
displays database information.
}

\cmdsum{\cmd{HELP}
   \param{option}
} {
displays information about the \caps{\PROGRAM} program.
}

\cmdsum{\cmd{LOG}
} {
requests that the log file be saved when the program is exited.
}

\cmdsum{\cmd{END}
} {
ends the equation input and begins the equation evaluation.
}

\cmdsum{\cmd{QUIT}
} {
ends the equation input and exits the program immediately without
writing an output database.
}
