\chapter{Site Supplements} \label{appx:site}

\section{VAX VMS}

The command to execute \caps{\PROGRAM} on \caps{VMS} is:
\begin{center}
\cmd{\PROGRAM\ \param{input\_database} \param{output\_database}
\param{user\_input}}
\end{center}

\param{Input\_database} is the filename of the input \exo{}
database. A prompt appears if \param{input\_database} is omitted. The
default is \cmd{TAPE11.EXO}.

\param{Output\_database} is the filename of the output \exo{}
database. A prompt appears if \param{output\_database} is omitted. The
default is \cmd{TAPE12.EXO}.

If \param{user\_input} is given, the user input is read from this
file. Otherwise user input is read from \caps{SYS\$INPUT} (the terminal
keyboard). User output is directed to \caps{SYS\$OUTPUT} (the terminal).

\caps{\PROGRAM} operates in either interactive or batch modes.

\section{CRAY CTSS}

To execute \caps{\PROGRAM}, the user must have selected the \cmd{acclib}
library and be running \cmd{ccl}.

The command to execute \caps{\PROGRAM} on \caps{CTSS} is:
\begin{center}
\cmd{algebra \param{input\_database} \param{output\_database}
   i=\param{input} o=\param{output}}
\end{center}

\param{Input\_database} is the filename of the input \exo{}
database. The default is \cmd{tape11}.

\param{Output\_database} is the filename of the output \exo{}
database. The default is \cmd{tape11}.

User input is read from \param{input}, which defaults to \cmd{tty} (the
terminal keyword). User output is directed to \param{output}, which
defaults to \cmd{tty} (the terminal).
