The commands are in free-format and must adhere to the following syntax
rules.
\setlength{\itemsep}{\medskipamount} \begin{itemize}
\item
Valid delimiters are a comma or one or more blanks.
\item
\ifx\PROGRAM\BLOT
Either lowercase or uppercase letters are acceptable, but lowercase
letters are converted to uppercase except in user-defined text that
appears on a plot (such as the plot caption).
\else
Either lowercase or uppercase letters are acceptable, but lowercase
letters are converted to uppercase.
\fi
\item
\ifx\PROGRAM\ALGEBRA
A ``\verb|'|'' character in any command line starts a comment. The
``\verb|'|'' and any characters following it on the line are ignored.
\else
A ``\cmd{\$}'' character in any command line starts a comment. The
``\cmd{\$}'' and any characters following it on the line are ignored.
\fi
\item
\ifx\PROGRAM\ALGEBRA
A command may be continued over several lines with an ``\verb|>|''
character. The ``\verb|>|'' and any characters following it on the
current line are ignored and the next line is appended to the current
line.
\else
A command may be continued over several lines with an ``\verb|>|''
character. The ``\verb|>|'' and any characters following it on the
current line are ignored and the next line is appended to the current
line.
\fi
\end{itemize}

Each command has an action keyword or ``verb'' followed by a variable
number of parameters.

The command verb is a character string. It may be abbreviated, as long
as enough characters are given to distinguish it from other commands.

The meaning and type of the parameters is dependent on the command verb.
Most command parameters are optional. If an optional parameter field is
blank, a command-dependent default value is supplied. Below is a
description of the valid entries for parameters.
\setlength{\itemsep}{\medskipamount} \begin{itemize}
%
\item
A numeric parameter may be a real number or an integer. A real number
may be in any legal \caps{FORTRAN} numeric format (e.g., \cmd{1},
\cmd{0.2}, \cmd{-1E-2}). An integer parameter may be in any legal
integer format.
\item
A string parameter is a literal character string. Most string parameters
may be abbreviated.
%
\newcommand{\okname}{f}
\ifx\PROGRAM\BLOT \renewcommand{\okname}{t} \fi
\ifx\PROGRAM\ALGEBRA \renewcommand{\okname}{t} \fi
\ifx\PROGRAM\GROPE \renewcommand{\okname}{t} \fi
\if\okname t
\item
Variable names must be fully specified. The blank delimiter creates a
problem with database variable names with embedded blanks. The program
handles this by deleting all embedded blanks from the input database
names. For example, the variable name ``\cmd{SIG~R}'' must be entered as
``\cmd{SIGR}''. The blank must be deleted in any references to the
variable.
\ifx\PROGRAM\GROPE
All database names appear exactly as input in all displays.
\else
All database names appear in uppercase without the embedded blanks in
all displays.
\fi
\fi
\ifx\PROGRAM\BLOT
\item
Screen and mesh positions may be selected with the graphics cursor (also
known as the graphics locator). Cursor input is device-dependent and
uses the VDI graphics locator routines. When the program prompts for the
position, the user positions the graphics cursor (e.g., the crosshairs)
on the screen, then selects the position by pressing any printable
keyboard character (e.g., the space bar).
\fi
\newcommand{\okrange}{f}
\ifx\PROGRAM\BLOT \renewcommand{\okrange}{t} \fi
\ifx\PROGRAM\GROPE \renewcommand{\okrange}{t} \fi
\if\okrange t
\ifx\PROGRAM\GROPE \newpage \fi %%%
\item
Several parameters allow a range of values. A range is in one of the
following forms:
\setlength{\itemsep}{\medskipamount} \begin{itemize}
\item ``\param{n$_{1}$}'' selects value \param{n$_{1}$},
\item ``\param{n$_{1}$} \cmd{TO} \param{n$_{2}$}'' selects all values from
\param{n$_{1}$} to \param{n$_{2}$},
\item ``\param{n$_{1}$} \cmd{TO} \param{n$_{2}$} \cmd{BY} \param{n$_{3}$}''
selects all values from \param{n$_{1}$} to \param{n$_{2}$} stepping by
\param{n$_{3}$}, where \param{n$_{3}$} may be positive or negative.
\end{itemize}
If the upper limit of the range is greater than the maximum allowable
value, the upper limit is changed to the maximum without a message.
\fi
\end{itemize}

\ifx\PROGRAM\ALGEBRA \newpage \fi %%%
The notation conventions used in the command descriptions are:
\setlength{\itemsep}{\medskipamount} \begin{itemize}
\item
The command verb is in \cmdverb{bold} type.
\item
A literal string is in all uppercase \cmd{SANSERIF} type and should be
entered as shown (or abbreviated).
\item
The value of a parameter is represented by the parameter name in
\param{italics}.
\newcommand{\okoptpar}{f}
\ifx\PROGRAM\BLOT \renewcommand{\okoptpar}{t} \fi
\ifx\PROGRAM\ALGEBRA \renewcommand{\okoptpar}{t} \fi
\ifx\PROGRAM\GROPE \renewcommand{\okoptpar}{t} \fi
\if\okoptpar t
\item
A literal string in square brackets (``[~]'') represents a parameter
option which is omitted entirely (including any following comma) if not
appropriate. These parameters are distinct from most parameters in that
they do not require a comma as a place holder to request the default
value.
\fi
\item
The default value of a parameter is in angle brackets (``$<$~$>$''). The
initial value of a parameter set by a command is usually the default
parameter value. If not, the initial setting is given with the default
or in the command description.
\end{itemize}
