\chapter{Command Input} \label{chap:command}

The user may direct the processing by entering commands to set
processing parameters.
\input{../include/cominput}

\cmddef{\cmdverb{DO4NOD}
   \cmd{ON} or \cmd{OFF} \default{\cmd{OFF}}
} {
\cmd{DO4NOD} indicates whether 4-node elements should be transformed into
four 4-node elements as is done with 8-node elements (if \cmd{ON}) or
left as 4-node elements (if \cmd{OFF}).
}

\cmddef{\cmdverb{EVARS}
   \param{record\_range$_{1}$}, \param{record\_range$_{2}$}, \ldots\
      \default{all element variable records}
} {
\cmd{EVARS} limits the selected element variable records. The element
variables are output only for selected record types. Element variable
records that are not selected are ignored.

The variable record types are listed in the \caps{ABAQUS} Manual
\cite{bib:abaqus}
Section 10.
Each \param{record\_range} must be in the range notation described above.
The effect of multiple \cmd{EVARS} commands is cumulative.
}

\cmddef{\cmdverb{NVARS}
   \param{record\_range$_{1}$}, \param{record\_range$_{2}$}, \ldots\
      \default{all nodal variable records}
} {
\cmd{NVARS} is the same as the \cmd{EVARS} command except that it
selects
nodal variable records.
}

\cmddef{\cmdverb{MATERIAL}
} {
\cmd{MATERIAL} signals the start of the element block settings. The
settings follow on the next lines. The \cmd{END} command ends the block
settings.

On each line following the \cmd{MATERIAL} command, the element block
number is given, followed by the numbers of the original elements that
are in that element block. The elements must be in the range notation
described above. All elements must be listed.

If no \cmd{MATERIAL} command is issued, the element block is determined
as described in Section~\ref{intro:element}.
}

\cmddef{\cmdverb{END}
} {
\cmd{END} ends the command input and starts the
\caps{ABAQUS} file translation.
}

\cmddef{\cmdverb{QUIT}
} {
\cmd{QUIT} ends the command input and aborts the program immediately
without writing an output database.
}
