\chapter{Command Input} \label{chap:command}

The user directs the processing by entering commands to set processing
parameters.
\input{texmisc:cominput}

\newpage
\section{Mesh Transformation}
\label{cmd:meshtrans}

\cmddef{\cmdverb{TRANSLATE}
   \param{ntran} \default{1},
   \param{tottran} \default{1.0},
   \param{grad} \default{1.0}
} {
\cmd{TRANSLATE} causes the \caps{2D} mesh to be translated to create the
\caps{3D} mesh. The number of levels is \param{ntran}, which is also the
number of \caps{3D} elements derived from each input \caps{2D} element.
The total range of the \caps{Z} coordinate is \param{tottran} with a
gradient of \param{grad}. This command negates a \cmd{ROTATE} command.

The gradient affects the spacing of the levels. The \caps{Z} range of
level \param{i} is
}

\cmddef{\cmdverb{ROTATE}
   \param{nrot} \default{1},
   \param{totdeg} \default{360.0},
   \param{cenrot} \default{0.0}
} {
\cmd{ROTATE} causes the \caps{2D} mesh to be rotated to create the
\caps{3D} mesh. The number of rotation levels is \param{nrot}, which is
also the number of \caps{3D} elements derived from each input \caps{2D}
element (with the exception of those affected by the \cmd{CENTER}
command). The mesh is rotated a total of \param{totdeg} degrees, with
each level rotated \param{totdeg}/\param{nrot} degrees. The center of
rotation \param{cenrot} is only meaningful if no center element blocks
are defined (see the \cmd{CENTER} command).
}

\cmddef{\cmdverb{REVOLVE}
   \param{axis$_{1}$}, \param{ndeg$_{1}$},
   \param{axis$_{2}$}, \param{ndeg$_{2}$}, \ldots\
      \default{last selection} \\
\cmdnext \cmdverb{REVOLVE}
   \cmd{RESET} \default{initial condition}
} {
\cmd{REVOLVE} causes the transformed \caps{3D} mesh to be rotated. Each
(\param{axis}, \param{ndeg}) parameter pair specifies an axis
(\cmd{X} or \cmd{Y} or \cmd{Z}) and the number of degrees to rotate. The
axis refers to the ``viewing'' axis, not to the object axis. The
rotations are according to right-hand rule. The center of the rotation
is specified by the \cmd{REVCEN} command.

Rotations are cumulative. The \cmd{REVOLVE RESET} command resets to no
rotation.
}

\cmddef{\cmdverb{REVCEN}
   \param{xcen} \default{\caps{2D} minimum X coordinate},
   \param{ycen} \default{\caps{2D} minimum Y coordinate},
   \param{zcen} \default{0.0}
} {
\cmd{REVCEN} sets the center of rotation for the \cmd{REVOLVE} command
to the point (\param{xcen},\param{ycen},\param{zcen}).
}

\newpage %%%
\cmddef{\cmdverb{OFFSET}
   \param{xoff} \default{0.0},
   \param{yoff} \default{0.0},
   \param{zoff} \default{0.0}
} {
\cmd{OFFSET} specifies offsets to be added to the transformed \caps{3D}
coordinates. If a \cmd{REVOLVE} command has been issued, the \caps{3D}
mesh is rotated before it is offset.
}

\newpage
\section{Element Block Types}
\label{cmd:blocktypes}

Each element block is assigned one of the following block types:
\setlength{\itemsep}{\smallskipamount} \begin{itemize}
\item A normal block requires no special handling.
\item A tunnel block (translation only) changes materials as it is
translated.
\item A center block (rotation around mesh edge only) has some elements
which border the mesh edge that is the center of rotation.
\end{itemize}

Initially, all blocks are normal blocks.

The \param{block\_id} parameter below refers to the element block identifier.
The identifiers are displayed by the \cmd{SHOW BLOCKS} command.

\cmddef{\cmdverb{BLOCK}
   \param{block\_id$_{1}$}, \param{block\_id$_{2}$}, \ldots\
      \default{all element blocks}
} {
\cmd{BLOCK} defines the specified element blocks
as normal blocks. This command
negates the effect of a \cmd{TUNNEL} or \cmd{CENTER} command.
}

\cmddef{\cmdverb{TUNNEL}
   \param{block\_id} \nodefault,
   \param{start} \default{1},
   \param{end} \default{number of levels},
   \param{inc} \default{1}
} {
\cmd{TUNNEL} defines the specified element block as a
tunnel block. A \cmd{TRANSLATE} command must be in effect before this
command is issued. If a \cmd{ROTATE} command is issued, all tunnel
blocks are changed to normal blocks.

For each tunnel block, a separate \caps{3D} element block is created
starting at level \param{start}, with each block having \param{inc}
levels. Any levels after level \param{end} are put in a single block.

For example, the commands
\cenlinesbegin
\cmd{TRANSLATE 15, 15.0} \\
\cmd{TUNNEL 999, 5, 9, 2}
\cenlinesend
create five blocks consisting of the following \caps{3D} elements
(derived from the \caps{2D} elements of element block \param{block\_id}):
\cenlinesbegin
1) the elements in levels 1, 2, 3, and 4, \\
2) the elements in levels 5 and 6, \\
3) the elements in levels 7 and 8, \\
4) the elements in level 9, \\
5) the elements in levels 10, 11, 12, 13, 14, and 15.
\cenlinesend

The block identifier of the first block is always \param{block\_id}. The new
blocks are assigned consecutive identifiers greater than the maximum
existing (and new) identifier.
}

\cmddef{\cmdverb{CENTER}
   \param{block\_id$_{1}$}, \param{block\_id$_{2}$}, \ldots\
      \default{all element blocks}
} {
\cmd{CENTER} defines the specified element blocks
as center blocks. A
\cmd{ROTATE} command must be in effect before this command is issued.
The mesh must be rotated a complete quadrant (90, 180, 270 or 360
degrees) and the number of rotation levels must be a multiple of 2 for
each 90 degrees of rotation. If a \cmd{TRANSLATE} command is issued, all
center blocks are changed to normal blocks.

If center blocks are defined, the center of rotation defined by the
\cmd{ROTATE} command is ignored. The center of rotation is the minimum
coordinate of all elements in the center blocks.
}

\newpage
\section{Front and Back Set Definition}
\label{cmd:setdef}

\cmddef{\cmdverb{NSETS}
   \cmd{FRONT} or \cmd{BACK} \nodefault,
   \param{set\_id$_{1}$}, \param{set\_id$_{2}$}, \ldots\ \nodefault
} {
\cmd{NSETS} defines front or back node sets with the given
identifiers. The identifiers must be unique from existing node
set identifiers and defined front and back node set identifiers.

Back sets cannot be defined on a 360-degree rotation.
}

\cmddef{\cmdverb{SSETS}
   \cmd{FRONT} or \cmd{BACK} \nodefault,
   \param{set\_id$_{1}$}, \param{set\_id$_{2}$}, \ldots\ \nodefault
} {
\cmd{SSETS} is equivalent to a \cmd{NSETS} command except that it
defines side sets.
}

\newpage
\section{Information and Processing}
\label{cmd:info}

\cmddef{\cmdverb{SHOW}
   \param{command} \default{no parameter}
} {
\cmd{SHOW} displays the settings of parameters relevant to the
\param{command}. For example, the command \cmd{SHOW BLOCK} displays
information about all the element blocks.
}

\cmddef{\cmdverb{LIST VARS}
} {
\cmd{LIST VARS} displays a summary of the input database. The summary
includes the database title; the number of nodes, elements, and
element blocks; and the number of node sets and side sets.
}

\cmddef{\cmdverb{HELP}
   \param{command} \default{no parameter}
} {
\cmd{HELP} displays information about the program command given as the
parameter. If no parameter is given, all the command verbs are
displayed. This command is system-dependent and may not be available on
some systems.
}

\cmddef{\cmdverb{END}
} {
\cmd{END} ends the command input and starts the database transformation.
}

\cmddef{\cmdverb{QUIT}
} {
\cmd{QUIT} ends the command input and exits the program immediately
without writing an output database.
}
